BOUT

### 2023-06-23 Data & Scripts/BOUT.Rmd
June 23rd
Late night, couldn't sleep :(

# BOUT

```{r Packages, include = F}
library("tidyverse")
library("psych")
library("car")
library("ggplot2")
library("ggpattern")
library("lmerTest")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
```

@@

This script will deal with analysing the relationship of the duration of bouts of behavior - head down, head up, and moving - with the presence of a sentinel, the presence of bait, the foraging environment and the group size.

As random effects, I've chosen the individual ID, the (julian) date, and the (decimal) time.

This analysis will answer the following question:\
Do individual vary the duration of behavioral bouts in response to their perception of risk and sentinel coverage in their foraging environment.

@@

```{r Crow Histogram, echo=F}
HIST<-ggplot(data = BOUT) +
  geom_histogram(aes(fill = BEHAVIOR, x = LDURATION)
                 , alpha = 1
                 , bins = 25)+
  scale_fill_manual(values=cbPalette
                    , labels = c("Foraging", "Alert", "Moving"))+
  xlab("Log Duration (s)")+
  ylab("Count")
HIST
```

@@

```{r BOUT MODEL 2}
BOUT.MOD2<-lmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)

@@

```{r RLMM MOD 2}
BOUT.RMOD2<-rlmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BEHAVIOR*GROUP_SIZE+BEHAVIOR*BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)

@@

The assumptions of the LMM do not appear to be violated.

@@

```{r Post Hoc}
DIFF<-emmeans::emmeans(BOUT.MOD2, ~SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+BAIT_PRESENCE,by = "BEHAVIOR", adjust = "fdr",pbkrtest.limit = 5070)
PAIRS<-emmeans::contrast(DIFF, "pairwise", adjust="fdr")
PAIRS
```

@@

```{r BxG.ENV BOUT Dot Plot}
BxG.ENV.DOT<-BOUT %>%
  summarySE(measurevar = "DURATION"
            , groupvars = c("BEHAVIOR", "GENERALIZED_ENVIRONMENT")) %>%
    ggplot(aes(x = BEHAVIOR
                     , y = DURATION
                     , color = BEHAVIOR))+
  geom_point(position = "dodge"
             , size = 3) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , guide="none")+
  scale_x_discrete(labels = c("Foraging", "Alert", "Moving"))+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()) +
  facet_grid(~GENERALIZED_ENVIRONMENT)

@@

GGPLOT! The error bars are not over the dots.

@@

@@

```{r Disturb}
DF.DOT<-BOUT %>%
    ggplot(aes(x = DISTURBANCE_FREQUENCY
               , y = DURATION
               , color = BEHAVIOR))+
  geom_point(position = "jitter"
             , size = 3) +
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , guide="none")+
  theme(axis.title.x = element_text("Disturbance per Minute")
        , legend.position = "bottom")

@@

>>>>

```{r Packages, include = F}
library("tidyverse")
library("psych")
library("car")
library("ggplot2")
library("ggpattern")
library("lmerTest")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
library("emmeans")
```

@@

This script will deal with analyzing the relationship of the duration of bouts of behavior - head down, head up, and moving - with the presence of a sentinel, the presence of bait, the foraging environment, the group size and the frequency of disturbances.

As random effects, I've chosen the video and individual ID

This analysis will answer the following question:\
Do individual vary the duration of behavioral bouts in response to their perception of risk and sentinel coverage in their foraging environment?

@@

```{r BOUT Histogram, echo=F}
BOUT.HIST<-ggplot(data = BOUT) +
  geom_histogram(aes(fill = BEHAVIOR, x = LDURATION)
                 , alpha = 1
                 , bins = 25)+
  scale_fill_manual(values=cbPalette
                    , labels = c("Foraging", "Alert", "Moving"))+
  xlab("Log Duration (s)")+
  ylab("Count")
BOUT.HIST
```

@@

```{r BOUT MODEL 2}
BOUT.MOD2<-lmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)

@@

```{r RLMM MOD 2}
BOUT.RMOD2<-rlmer(LDURATION~BEHAVIOR*SENTINEL_PRESENCE+BEHAVIOR*GENERALIZED_ENVIRONMENT+SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT+GROUP_SIZE+BAIT_PRESENCE+DISTURBANCE_FREQUENCY+(1|VIDEO_ID/ID), data = BOUT)

@@

There's a horizontal line on the Q-Q vs. random effects. In the outputs of the models, the videoID alone does not account for any variation.

@@

```{r BOUT Post Hoc}
BOUT.DIFF<-emmeans(BOUT.MOD2, ~BEHAVIOR*SENTINEL_PRESENCE*GENERALIZED_ENVIRONMENT, pbkrtest.limit = 5070)
test(pairs(BOUT.DIFF, by= "BEHAVIOR"), adjust="fdr")
```
Very cool! These are the results of pairwise comparisons averaged over the levels of group size and bait presence (**Not disturbance frequency**).

These pairwise comparisons show there are no significant differences in the bout duration of 'moving' or 'head up' between each combination of 'sentinel_presence' and 'generalized_environment'.

However, there are differences in the bout duration of 'head down' behavior. There are differences caused by the presence of a sentinel in commercial areas. There are significant differences between bout durations in commercial areas without a sentinel and bout durations in green areas, in both the presence and absence of a sentinel.

Let's quickly have a look at that:

```{r Pairwise comparisons}
BOUT.DOT
```

Huh... I would have expected that there would be significant differences in bouts of 'head up' behavior.

@@

```{r BxG.ENV BOUT Dot Plot}
BxG.ENV.DOT<-BOUT %>%
  summarySE(measurevar = "DURATION"
            , groupvars = c("BEHAVIOR", "GENERALIZED_ENVIRONMENT")) %>%
    ggplot(aes(x = BEHAVIOR
                     , y = DURATION
                     , color = BEHAVIOR))+
  geom_point(position = position_dodge(width=0.9)
             , size = 3) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.1
                , position = position_dodge(width=0.9))+
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , guide="none")+
  scale_x_discrete(labels = c("Foraging", "Alert", "Moving"))+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()) +
  facet_grid(~GENERALIZED_ENVIRONMENT)

@@

Ok, we see that smaller group sizes lead to shorter bout durations.

@@

The presence of bait has the effect of decreasing the duration of 'head down' and 'head up' behavior, yet increased the duration of bouts of movement.

@@

```{r Disturb}
DF.DOT<-BOUT %>%
    ggplot(aes(x = DISTURBANCE_FREQUENCY
               , y = DURATION
               , color = BEHAVIOR))+
  geom_point(size = 3) +
  theme_classic() +
  labs(x="Mean Bout Duration (s)", y="Disturbances per Minute", color= "Behavior") +
  scale_colour_manual(values = cbPalette, labels = c("Foraging", "Alert", "Moving"))+
  scale_x_continuous(n.breaks=14)
  theme(legend.position = "bottom")

@@

Finally, it seems like as the frequency of disturbances increases, the duration of bouts decreases.
